Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
changed div ex notation
flip order of monomials in block order def & g0 ex

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \di X \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

\[
	\di \sx = \di X \oplus \left(\bigoplus_{i
	= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X
\]
\end{defn}

@@

\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d f_i  > \sum_{i = 1}^{n} d f_i' =: \deg \beta
\end{equation}
or $\alpha$ is larger using the lexicographical ordering event of equal
degrees:

@@

\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d f_i  > \sum_{i = 1}^{n} d f_i' =: \deg \beta
\end{equation}
or $\alpha$ has a smaller right-most entry in the event of equal
degrees:

@@

\begin{defn}
\label{defn:block-order}
\todo{Aaron: Doesn't \cite{vzb:stacky} put the $y_i$ before the $x_i$? Shouldn't we match this convention (here and anywhere we use this in the paper later?}
Let $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$ and $k[y_{c_1, 1}, \ldots,
y_{c_m, m}]$ be weighted polynomial rings with existing term
orders $\prec_x$ and $\prec_y$, respectively. Let $\alpha :=
\prod_{i = 1}^{n} x_{d_i, i}^{f_i}$ $\prod_{j = 1}^{m} y_{c_j, j}
^{h_i}$ and $\beta := \prod_{i = 1}^{n} x_{d_i, i}^{f_i'}
\prod_{j = 1}^{m} y_{c_j, j}^{h_i'}$ be monomials in $k[x_{d_1, 1},
\ldots, x_{d_n, n}] \otimes$ $k[y_{d_1', 1}, \ldots, y_{d_m', m}]$.
Let $\alpha_x := \prod_{i = 1}^{n} x_{d_i, i}^{f_i}$ be the
part of $\alpha$ in $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$ and
likewise with $\alpha_y$, $\beta_x$, and $\beta_y$.

In the \textbf{(graded) block} (or \textbf{elimination}) term
ordering on $k[x_{d_1, 1}, \ldots, x_{d_n, n}] \otimes k[y_{d_1', 1},
\ldots, y_{d_m', m}]$, we define $\alpha \succ \beta$ if
  \begin{enumerate}
	\item[(i)] $\deg \alpha  > \deg \beta$ or
  \item[(ii)] $\deg \alpha = \deg \beta$ and $\alpha_y \succ_y \beta_y$ or
	\item[(iii)] $\deg \alpha = \deg \beta$ and $\alpha_y = \beta_y$ and $\alpha_x \succ_x \beta_x$
  \end{enumerate}
\end{defn}

@@

We can directly see that the relations are given in degrees $10$
and $14$. In particular, we have relations
\begin{align*}
	&a_1 x_{3, 1} x_{7, 2} + a_2 x_{3, 1} x_{7, 1} + a_3 x_{5, 1}^2 = 
	0 &\text{ in degree $10$} \\
	&b_1 x_{7, 2}^2 + b_2 x_{7, 1} x_{7, 2} + b_3 x_{7, 1}^2
	+ b_4 x_{3, 1}^3 x_{5, 1} = 0  &\text{ in degree $14$}
\end{align*}

\noindent
and note that $a_1$ and $b_1$ are both nonzero. If $a_1 = 0$, then we
have that
\[
	a_2 x_{3, 1} x_{7, 1} = -a_3 x_{5, 1}^2
\]

\noindent
However, $x_{3, 1} x_{7, 1}$ has pole orders $(3, 4, 3)$ and
$x_{5, 1}^2$ has pole orders $(2, 4, 4)$ so their
coefficients must be zero for this equality to hold. We have a
similar pole order consideration in forcing $b_1$ to be nonzero.

Let $I$ be the ideal generated by these relations in
$k[x_{7, 1}, x_{7, 2}, x_{5, 1}, x_{3, 1}]$. Under grlex with
$x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$, the initial
ideal of $I$ is
\[
	\initial(I) = \langle x_{3, 1} x_{7, 2}, x_{7, 2}^2 \rangle
\]

@@

To demonstrate that these two relations generate all 
relations, we show that $\dim_k (R')_d = \dim_k
(k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / \initial(I))_d$ for 
all degrees $d \geq 0$. In particular,

\begin{align*}
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / I)_d &=
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}]	/ in_\prec(I))_d \\
	&= \dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / \langle
	x_{3, 1} x_{7, 2}, x_{7, 2}^2 \rangle
\end{align*}

\noindent
for all $d \geq 0$. But the fact $\dim_k (R')_d = \dim_k (k[x_{3, 1},
x_{5, 1}, x_{7, 1}, x_{7, 2}] / \langle x_{5, 1}^2,	x_{7, 1}^2 \rangle$ can
be checked by looking at the $k$-basis at each degree $d$.

Therefore, the canonical ring $R'$ has presentation $R' =
k[x_{7, 1}, x_{7, 2}, x_{5, 1}, x_{3, 1}] / I$ with
initial ideal $\initial(I)$ generated by quadratics under grlex
with $x_{7,1} \prec x_{7,2} \prec x_{5,1} \prec x_{3,1}$.
%These quadratics must also determine minimal generators for $I$ by
%Lemma ~\ref{lem:minimal_quadratic}.
Thus, $R'$ is generated up to degree $e = 7$ with relations up to
degree $2e = 14$, as desired.

@@

\begin{enumerate}
	\item In the case $g \geq 2$, Proposition ~\ref{prop:relation_11} gives a bound on the degree of generators of the relations of a log spin canonical ring $R_L$. Additionally, by Remark ~\ref{rem:relations_generation_ten}, any curve of genus 2 with half canonical divisor $L = P$ has spin canonical ring $R_L$ with relations in degree 10. This raises the question of whether the bound of 11 on the degree of the relations in $R_L$ can be reduced to $10$. In particlar, we make the following conjecture.
\begin{conjecture}
\label{conj:relations-10}
If $X$ is a curve of genus $g \geq 2$ and $L \in \di X$ with $2L \sim K,$ the ideal of relations in $R_L$ is generated in degree at most $10$. More precisely, if $x_1,\ldots, x_n$ is a minimal set of generators for $R_L$ and $\phi:k[y_1,\ldots, y_n] \rightarrow R_L, y_i \mapsto x_i,$ then $I_L := \ker \phi$, is generated in degree at most 10. 
\end{conjecture}
Note that in the case $\dim L \geq 2$ and $L$ is basepoint-free, by Neves ~\cite[Proposition III.12]{neves:halfcan}, $I_L$ is generated in degree at most 8. Additionally, if Conjecture ~\ref{conj:relations-10} were true, by Theorem ~\ref{thm:main} we would immediately obtain the following corollary, providing a more concise statement than that of Theorem ~\ref{thm:main}.
\begin{cor}
\label{cor:main-refined}
Let $(\sx,\Delta,L)$ be a log spin curve over a perfect field $k$ with signature $(g;e_1,\ldots, e_r;\delta)$ and let $R_L$ be the log canonical ring. Suppose Conjecture ~\ref{conj:relations-10}. Then, $R_L$ is generated in degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e,$ apart from a finite list of exceptions.
\end{cor}
	\item Another direction for further research is to extend the results of this paper to divisors $D \in \di \sx$ on a stacky curve $\sx,$ where $nD \sim K.$ The canonical rings of such divisors would translate to rings of fractional weight modular forms. As defined in Adler and Ramanan, immediately following \cite[Corollary 24.5]{adler:moduli}, a {\bf modular form of fractional weight} $\frac{2a}{n}$ on $X(p)$ is a section of a line bundle $\eta$ so that there exists a line bundle $\gamma$ with $\gamma^{\otimes a} \cong \eta$ and $\gamma^{n} \cong \Omega_{X(p)}^1$. Adler and Ramanan \cite[Appendix 2]{adler:moduli} construct an explicit basis of weight $\frac{4}{5}$ modular forms for $X(11)$. Fractional modular forms also make an apperance in Milnor \cite[$\mathsection$ 6]{milnor:fractional-weight}. In this work, Milnor shows that for a triangle group $\Gamma$ of signature $(0; e_1,e_2,e_3; 0)$ there exist generators $f_1,f_2,f_3$ of a certain ring of fractional weight modular forms, satisfying $f_1^{e_1}+f_2^{e_2}+f_3^{e_3} = 0$.

>>>>

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle P  \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle P  \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\end{defn}

@@

\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d_i f_i  > \sum_{i = 1}^{n} d_i f_i' =: \deg \beta
\end{equation}
or $\alpha$ is larger using the lexicographical ordering event of equal
degrees:

@@

\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d_i f_i  > \sum_{i = 1}^{n} d_i f_i' =: \deg \beta
\end{equation}
or $\alpha$ has a smaller right-most entry in the event of equal
degrees:

@@

\begin{defn}
\label{defn:block-order}
Let $k[y_{c_1, 1}, \ldots, y_{c_m, m}]$ and $k[x_{d_1, 1}, \ldots,
x_{d_n, n}]$ be weighted polynomial rings with existing term
orders $\prec_y$ and $\prec_x$, respectively. Let $\alpha :=
\prod_{j = 1}^{m} y_{c_j, j}^{h_i}$ $\prod_{i = 1}^{n} x_{d_i, i}
^{f_i}$ and $\beta := \prod_{j = 1}^{m} y_{c_j, j}^{h_i'}
\prod_{i = 1}^{n} x_{d_i, i}^{f_i'}$ be monomials in $k[y_{d_1', 1},
\ldots, y_{d_m', m}] \otimes$ $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$.
Let $\alpha_y := \prod_{j = 1}^{m} y_{c_j, j}^{h_i}$ be the
part of $\alpha$ in $k[y_{c_1, 1}, \ldots, y_{c_m, m}]$ and
likewise with $\alpha_x$, $\beta_y$, and $\beta_x$.

In the \textbf{(graded) block} (or \textbf{elimination}) term
ordering on $k[y_{d_1', 1}, \ldots, y_{d_m', m}] \otimes k[x_{d_1, 1},
\ldots, x_{d_n, n}]$, we define $\alpha \succ \beta$ if
  \begin{enumerate}
	\item[(i)] $\deg \alpha  > \deg \beta$ or
  \item[(ii)] $\deg \alpha = \deg \beta$ and $\alpha_y \succ_y \beta_y$ or
	\item[(iii)] $\deg \alpha = \deg \beta$ and $\alpha_y = \beta_y$ and $\alpha_x \succ_x \beta_x$
  \end{enumerate}
\end{defn}

@@

We can directly see that the relations are given in degrees $10$
and $14$. In particular, we have relations
\begin{align*}
	&a_1 x_{7, 2} x_{3, 1} + a_2 x_{7, 1} x_{3, 1} + a_3 x_{5, 1}^2 = 
	0 &\text{ in degree $10$} \\
	&b_1 x_{7, 2}^2 + b_2 x_{7, 2} x_{7, 1} + b_3 x_{7, 1}^2
	+ b_4 x_{5, 1} x_{3, 1}^3 = 0  &\text{ in degree $14$}
\end{align*}

\noindent
and note that $a_1$ and $b_1$ are both nonzero. If $a_1 = 0$, then we
have that
\[
	a_2 x_{7, 1} x_{3, 1} = -a_3 x_{5, 1}^2
\]

\noindent
However, $x_{7, 1} x_{3, 1}$ has pole orders $(3, 4, 3)$ and
$x_{5, 1}^2$ has pole orders $(2, 4, 4)$ so their
coefficients must be zero for this equality to hold. We have a
similar pole order consideration in forcing $b_1$ to be nonzero.

Let $I$ be the ideal generated by these relations in
$k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]$. Under grlex with
$x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$, the initial
ideal of $I$ is
\[
	\initial(I) = \langle x_{7, 2}^2, x_{7, 2} x_{3, 1} \rangle
\]

@@

To demonstrate that these two relations generate all 
relations, we show that $\dim_k (R')_d = \dim_k
(k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \initial(I))_d$ for 
all degrees $d \geq 0$. In particular,

\begin{align*}
	\dim_k (k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / I)_d &=
	\dim_k (k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}]	/ in_\prec(I))_d \\
	&= \dim_k (k[x_{7, 2}, x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle
	x_{7, 2}^2, x_{3, 1} x_{7, 2} \rangle
\end{align*}

\noindent
for all $d \geq 0$. But the fact $\dim_k (R')_d = \dim_k (k[x_{7, 2},
x_{7, 1}, x_{5, 1}, x_{3, 1}] / \langle x_{7, 2}^2, x_{3, 1} x_{7, 2}
\rangle$ can be checked by looking at the $k$-basis at each degree
$d$.

Therefore, the canonical ring $R'$ has presentation $R' =
k[x_{7, 2}^2, x_{3, 1} x_{7, 2}] / I$ with
initial ideal $\initial(I)$ generated by quadratics under grlex
with $x_{3,1} \prec x_{5,1} \prec x_{7,1} \prec x_{7,2}$.
%These quadratics must also determine minimal generators for $I$ by
%Lemma ~\ref{lem:minimal_quadratic}.
Thus, $R'$ is generated up to degree $e = 7$ with relations up to
degree $2e = 14$, as desired.

@@

\begin{enumerate}
	\item In the case $g \geq 2$, Proposition ~\ref{prop:relation_11} gives a bound on the degree of generators of the relations of a log spin canonical ring $R_L$. Additionally, by Remark ~\ref{rem:relations_generation_ten}, any curve of genus 2 with half canonical divisor $L = P$ has spin canonical ring $R_L$ with relations in degree 10. This raises the question of whether the bound of 11 on the degree of the relations in $R_L$ can be reduced to $10$. In particular, we make the following conjecture.
\begin{conjecture}
\label{conj:relations-10}
If $X$ is a curve of genus $g \geq 2$ and $L \in \di X$ with $2L \sim K,$ the ideal of relations in $R_L$ is generated in degree at most $10$. More precisely, if $x_1,\ldots, x_n$ is a minimal set of generators for $R_L$ and $\phi:k[y_1,\ldots, y_n] \rightarrow R_L, y_i \mapsto x_i,$ then $I_L := \ker \phi$, is generated in degree at most 10. 
\end{conjecture}
Note that in the case $\dim L \geq 2$ and $L$ is basepoint-free, by Neves ~\cite[Proposition III.12]{neves:halfcan}, $I_L$ is generated in degree at most 8. Additionally, if Conjecture ~\ref{conj:relations-10} were true, by Theorem ~\ref{thm:main} we would immediately obtain the following corollary, providing a more concise statement than that of Theorem ~\ref{thm:main}.
\begin{cor}
\label{cor:main-refined}
Let $(\sx,\Delta,L)$ be a log spin curve over a perfect field $k$ with signature $(g;e_1,\ldots, e_r;\delta)$ and let $R_L$ be the log canonical ring. Suppose Conjecture ~\ref{conj:relations-10}. Then, $R_L$ is generated in degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e,$ apart from a finite list of exceptions.
\end{cor}
	\item Another direction for further research is to extend the results of this paper to divisors $D \in \di \sx$ on a stacky curve $\sx,$ where $nD \sim K.$ The canonical rings of such divisors would translate to rings of fractional weight modular forms. As defined in Adler and Ramanan, immediately following \cite[Corollary 24.5]{adler:moduli}, a {\bf modular form of fractional weight} $\frac{2a}{n}$ on $X(p)$ is a section of a line bundle $\eta$ so that there exists a line bundle $\gamma$ with $\gamma^{\otimes a} \cong \eta$ and $\gamma^{n} \cong \Omega_{X(p)}^1$. Adler and Ramanan \cite[Appendix 2]{adler:moduli} construct an explicit basis of weight $\frac{4}{5}$ modular forms for $X(11)$. Fractional modular forms also make an apperance in Milnor \cite[$\mathsection$ 6]{milnor:fractional-weight}. In this work, Milnor shows that for a triangle group $\Gamma$ of signature $(0; e_1,e_2,e_3; 0)$ there exist generators $f_1,f_2,f_3$ of a certain ring of fractional weight modular forms, satisfying $f_1^{e_1}+f_2^{e_2}+f_3^{e_3} = 0$.
